Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-06 ms.tex
section 5.5.1

The total gas heating rate, 
\begin{equation}
\vwO = \sqrt{(v_{w}^{\star})^{2} + (v_{w}^{\rm MSP})^{2} + (v_{w}^{\rm Ia})^{2} + (v_{w}^{\bullet})^{2}},
\label{eq:vtot}
\end{equation}
includes contributions from stellar winds, supernovae, pulsars, and radiative SMBH feedback.  The strength of each heating source depends explicitly on the SMBH mass and the stellar population in the galactic nuclear region.  The latter could best be described by a single star burst formation in the past, or by a more continuous star formation history that itself varies systematically with the galaxy mass and hence $\Mbh$.

@@

We crudely account for the expected suppression of SN Ia heating
resulting from its non-steady nature by setting $v_{w}^{\rm Ia} = 0$
when the stagnation radius (calculated excluding Ia heating) exceeds
the Ia radius $r_{\rm Ia}$ exterior to which the interval between
subsequent supernovae exceeds the local dynamical time
(eq.~[\ref{eq:rIa}]) {\bf AG have not yet done this in the impulsive limit}.  Finally, note that $v_{\rm w}^{\rm C}$ and
depends on the total heating rate (eq.~[\ref{eq:vtot}]), requiring us
to simultaneously solve a series of implicit equations to determine
each heating rate.  Heating from MSPs is found to be negligibly small
for fiducial parameters and hence is not shown, while kinetic feedback
is neglected given its uncertain efficiency ($\S\ref{sec:kinetic}$).

For low-mass black holes, stellar winds are the most important source
of heating on timescales of $\tau_{\star} \lesssim $ few $10^{7}$
years.  Compton heating comes to dominate at later times due to the
higher accretion rates that accompany the overall decrease in the
heating rate, coupled with the persistently high mass loss rates and
gas densities associated with a relatively young stellar population.
Compton heating dominates at all epochs for higher mass black holes,
with stellar winds and Ia SN competing for the most significant
non-AGN heating source.  As the total heating rate declines with time,
after some point the flow becomes thermally unstable according to the
criterion $(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|)_{r_{s}} \lesssim
10$ (eq.~[\ref{eq:cooling3}]) {\bf BDM: correct?}, as shown by dashed
lines in Fig.~\ref{fig:vwSourcesImp}.  The thermal stability threshold
is calculated neglecting Compton heating because SMBH feedback may not
be capable of truly stabilizing the flow given its inability to
respond instantaneously to local changes in the properties of the gas.
Thermally unstable flow is seen to be the norm in the single starburst
case, except at very early times, $\tau_{\star} \lesssim 1-3 \times
10^7$ yr.

>>>>

The total external gas heating rate, 
\begin{equation}
\vwO = \sqrt{(v_{w}^{\star})^{2} + (v_{w}^{\rm MSP})^{2} + (v_{w}^{\rm Ia})^{2} + (v_{w}^{\bullet})^{2}},
\label{eq:vtot}
\end{equation}
includes contributions from stellar winds, supernovae, pulsars, and radiative SMBH feedback.  The strength of each heating source depends explicitly on the SMBH mass and the stellar population in the galactic nuclear region.  As we now describe, the latter could best be described by a single star burst formation in the past, or by a more continuous star formation history that itself varies systematically with the galaxy mass and hence $\Mbh$.

@@

We crudely account for the expected suppression of SN Ia heating
resulting from its non-steady nature by setting $v_{w}^{\rm Ia} = 0$
when the stagnation radius (calculated excluding Ia heating) exceeds
the Ia radius $r_{\rm Ia}$ exterior to which the interval between
subsequent supernovae exceeds the local dynamical time
(eq.~[\ref{eq:rIa}]).  Finally, note that $v_{\rm w}^{\rm C}$ and
depends on the total heating rate (eq.~[\ref{eq:vtot}]), requiring us
to simultaneously solve a series of implicit equations to determine
each heating rate.  Heating from MSPs is found to be negligibly small
for fiducial parameters and hence is not shown, while kinetic feedback
is neglected given its uncertain efficiency ($\S\ref{sec:kinetic}$).

For low-mass black holes (Fig.~\ref{fig:vwSourcesImp}, top panel), stellar winds are the most important source
of heating on timescales of $\tau_{\star} \lesssim 10^{7}$
years.  Compton heating comes to dominate at later times due to the
higher accretion rates that accompany the overall decrease in the
heating rate, coupled with the persistently high mass loss rates and
gas densities associated with a relatively young stellar population.  Interestingly, Ia SN dominate the heating briefly around 10$^{9}$ yr due to a window of time in which $r_Ia < \rs$.  For high-mass black holes (Fig.~\ref{fig:vwSourcesImp}, bottom panel), Compton heating dominates at all epochs, while SN Ia  dominate after 10$^{8}$ yr.

As the total heating rate declines with time, after some point the flow becomes thermally unstable according to the
criterion $(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|)_{r_{s}} \lesssim 10$ (eq.~[\ref{eq:cooling3}]), as shown by dashed
lines in Fig.~\ref{fig:vwSourcesImp}.  Compton heating is neglected in calculating thermal stability because unlike local stellar feedback mechanisms, SMBH feedback may not be capable of truly stabilizing the flow given its inability to
respond instantaneously to local changes in the properties of the gas.  Thermally unstable flow is present throughout the single starburst case except at very early times, $\tau_{\star} \lesssim 10^7$ yr, and at late times $\tau_{\star} \gtrsim 10^{8}$ yr in the high $M_{\bullet}$ case.
